The low-frequency Murchison radio environment:\\interference analysis and mitigation

### 2014-09-28 articles/2014/MWA-Radio-Environment/mwa-radio-environment.tex
Finishing log N log S plots with new data

\section{Method}
RFI detection, often referred to as ``data flagging'', is one of the first steps in the data processing. It is important to perform initial data flagging at high time and frequency resolution, because this increases detection accuracy and decreases the loss of data \citep{lofar-radio-environment}. Consequently, RFI detection has to work on large data volumes, and its computational cost is therefore a concern -- in particular for a many-element arrays. In some projects, simple thresholding is used to mitigate the worst interference, which is computationally cheap but not very accurate. For example, a $3\sigma$ threshold is used for reducing PAPER data in \citet{parsons-paper-eorlimit-2014}. Several observatories or projects have designed pipelines that include more advanced RFI mitigation. Examples of such pipelines include \textsc{aoflagger} \citep{post-correlation-rfi-classification,scale-invariant-rank-operator}, originally designed for LOFAR; \textsc{flagcal} for preprocessing data from the Giant Metrewave Radio Telescope (GMRT; \citealt{prasad-flagcal-2012}); and \textsc{serpent} for preprocessing data from the Multi-Element Radio Linked Interferometer Network (e-MERLIN; \citealt{serpent-peck-2013}). LOFAR's \textsc{aoflagger} has shown to have a good accuracy, is fast and has a library interface, which allows it to be integrated in a pipeline. It is therefore used for MWA's RFI detection.

@@

Both projects avoid the subbands that cover the frequency range 133.8--138.9~MHz, because the ORBCOMM low Earth orbitting (LEO) satellites transmit in these subbands. When these frequencies are observed, the subbands are often so strongly contaminated that it affects imaging sensitivity. For completeness, we include an 16-min observation that covers this frequency range.

>>>>

\section{Method}
RFI detection, often referred to as ``data flagging'', is one of the first steps in the data processing. It is important to perform initial data flagging at high time and frequency resolution, because this increases detection accuracy and decreases the loss of data \citep{lofar-radio-environment}. Consequently, RFI detection has to work on large data volumes, and its computational cost is therefore a concern -- in particular for a many-element arrays. In some projects, simple thresholding is used to mitigate the worst interference, which is computationally cheap but not very accurate. For example, a $3\sigma$ threshold is used for reducing PAPER data in \citet{parsons-paper-eorlimit-2014}. Several observatories or projects have designed pipelines that include more advanced RFI mitigation. Examples of such pipelines include \textsc{aoflagger} \citep{post-correlation-rfi-classification,scale-invariant-rank-operator}, originally designed for LOFAR; \textsc{flagcal} for preprocessing data from the Giant Metrewave Radio Telescope (GMRT; \citealt{prasad-flagcal-2012}); and \textsc{serpent} for preprocessing data from the Multi-Element Radio Linked Interferometer Network (e-MERLIN; \citealt{serpent-peck-2013}). For RFI detection in MWA observations, LOFAR's \textsc{aoflagger} is used. This flagger has shown to have a good accuracy, is fast and has a library interface, which allows it to be integrated in a pipeline.

@@

Both projects avoid the subbands that cover the frequency range 133.8--138.9~MHz, because the ORBCOMM low Earth orbitting (LEO) satellites transmit in these subbands. When these frequencies are observed, the subbands are often so strongly contaminated that it affects imaging sensitivity. For completeness, we include a 16-min observation that covers this frequency range.
